---
title: "Minimal BC-GPSM Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimal BC-GPSM Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

This article shows a minimal end-to-end workflow for `BC.GPSM` (`BC-GPSM`). The example
uses simulated data with three treatment groups, six baseline covariates, and a
continuous outcome.

```{r setup, include=FALSE}
if (identical(tolower(Sys.getenv("IN_PKGDOWN")), "true")) {
  r_dir <- if (dir.exists("R")) "R" else file.path("..", "R")
  if (dir.exists(r_dir)) {
    invisible(lapply(list.files(r_dir, pattern = "[.]R$", full.names = TRUE), source))
  }
} else {
  library(BC.GPSM)
}
```

```{r load-package, eval=FALSE}
library(BC.GPSM)
```

## Simulate Example Data

```{r simulate-data}
set.seed(123)
n <- 300
z1 <- rnorm(n)
z2 <- rnorm(n)
age <- 50 + 8 * z1
biomarker <- 0.6 * z2 + rnorm(n, sd = 0.5)
severity <- 0.35 * z1 - 0.25 * z2 + rnorm(n, sd = 0.9)
distance <- runif(n, -1, 1)
prior_tx <- rbinom(n, 1, plogis(0.1 * z1 - 0.1 * z2))
clinic <- factor(rbinom(n, 1, 0.45), labels = c("Rural", "Urban"))

score <- cbind(
  A = 0,
  B = 0.35 * (0.05 * (age - 50) - 0.20 * biomarker +
    0.18 * severity + 0.20 * prior_tx),
  C = 0.35 * (-0.02 * (age - 50) + 0.25 * biomarker -
    0.18 * distance + 0.18 * (clinic == "Urban"))
)
prob <- exp(score) / rowSums(exp(score))
trt <- apply(prob, 1, function(p) sample(c("A", "B", "C"), 1, prob = p))

dat <- data.frame(
  trt = factor(trt, levels = c("A", "B", "C")),
  age = age,
  biomarker = biomarker,
  severity = severity,
  distance = distance,
  prior_tx = prior_tx,
  clinic = clinic
)
dat$y <- 1 + 0.5 * (dat$trt == "B") + 1.0 * (dat$trt == "C") +
  0.03 * age - 0.4 * biomarker + 0.25 * severity - 0.2 * distance +
  0.3 * prior_tx + 0.2 * (clinic == "Urban") + rnorm(n)

head(dat)
```

## Fit the Main Estimator

Use `dr_gpsm()` for the cross-fitted estimator. In this example, treatment
`A` is used as the reference group, so contrast labels such as `BvA` mean
the estimated mean potential outcome under `B` minus the estimated mean
potential outcome under `A`.

```{r fit-dr-gpsm}
fit <- dr_gpsm(
  data = dat,
  treatment = 1,
  treatment_ref = "A",
  covariate = 2:7,
  outcome = 8,
  gps_model = "logit",
  outcome_model = "lm",
  folds = 2,
  nboot = 50
)

data.frame(
  contrast = names(fit$estimate),
  estimate = fit$estimate,
  ci_lower = fit$ci_lower,
  ci_upper = fit$ci_upper,
  row.names = NULL
)
```

For final analyses, use a larger `nboot` such as `500` or more. The smaller
value above keeps the example quick to run.

## Inspect GPS Overlap

`gps_pre_process()` can fit a full-sample diagnostic GPS model and add `gps_`
and `loggps_` columns. These diagnostic scores are useful for visualizing
overlap; `dr_gpsm()` itself uses cross-fitted GPS values internally.

```{r gps-diagnostic}
gps_dat <- gps_pre_process(
  data = dat,
  treatment = 1,
  treatment_ref = "A",
  covariate = 2:7,
  gps_model = "logit"
)

gps_histogram(gps_dat, bins = 20)
```

## Check Covariate Balance

Use `balance_check_plot()` to compare pairwise covariate balance before and
after nearest-neighbor matching. The plot reports absolute standardized mean
differences; smaller values indicate better balance. The example standardizes
covariates before matching so that variables on different scales contribute
comparably to the matching distance.

```{r balance-check}
balance_check_plot(
  data = dat,
  treatment = 1,
  treatment_ref = "A",
  covariate = 2:7,
  match_on = "covariates",
  standardize = TRUE,
  style = "love"
)
```

The same function can also use existing `loggps_*` columns. For example, after
creating `gps_dat` above:

```{r balance-check-gps, eval=FALSE}
balance_check_plot(
  data = gps_dat,
  treatment = 1,
  treatment_ref = "A",
  covariate = 2:7,
  match_on = "gps",
  fit_gps = FALSE
)
```

## Choosing Options

Start with transparent models:

- `gps_model = "logit"` for multinomial logistic GPS estimation;
- `outcome_model = "lm"` for a simple outcome-regression adjustment;
- `treatment_ref` set explicitly when a natural control group exists.

Then compare sensitivity to more flexible nuisance models such as `gbm`,
`rf`, or `gam` when the sample size is large enough to support them.

## Optional Learners and Tuning

The baseline `logit + lm` workflow does not require the optional modeling
stack. Install only the packages needed by the selected model or tuning path:

```{r optional-packages, eval=FALSE}
install.packages(c(
  "gbm", "mgcv", "VGAM", "randomForest", "glmnet",
  "caret", "ModelMetrics", "pROC", "foreach"
))
```

`xgboost` and `ranger` provide additional GPS and outcome-model backends:

```{r experimental-packages, eval=FALSE}
install.packages(c("xgboost", "ranger"))
```

These two backends are experimental and remain under active testing. For now,
use them as sensitivity analyses and report their package versions and model
settings whenever results are shared.
